---
title: "Imputation"
author: "Ishaan Lodhi"
date: "2024-11-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Enable all packages here
library(tidyverse)
library(arrow)
library(imputeTS)
```

```{r}
#File 1
file_1 <- read_parquet("/Users/ishaan/Downloads/part-1.parquet")
col_drop <- c("in.city", "in.clothes_washer_presence", "in.county_and_puma", "in.federal_poverty_level", "in.geometry_floor_area_bin", "in.geometry_stories_low_rise", "in.income_recs_2015", "in.income_recs_2020", "in.puma", "in.puma_metro_status", "in.vintage")
file_1 <- file_1[, !colnames(file_1) %in% col_drop]
head(file_1)
gc()
```

```{r}
file_2 <- read_parquet("/Users/ishaan/Downloads/part-2.parquet")
file_2 <- file_2[, !colnames(file_2) %in% col_drop]
gc()
```

```{r}
file_3 <- read_parquet("/Users/ishaan/Downloads/part-3.parquet")
file_3 <- file_3[, !colnames(file_3) %in% col_drop]
gc()
```

```{r}
file_4 <- read_parquet("/Users/ishaan/Downloads/part-4.parquet")
file_4 <- file_4[, !colnames(file_4) %in% col_drop]
gc()
```

```{r}
file_5 <- read_parquet("/Users/ishaan/Downloads/part-5.parquet")
file_5 <- file_5[, !colnames(file_5) %in% col_drop]
gc()
```

```{r}
file_6 <- read_parquet("/Users/ishaan/Downloads/part-6.parquet")
file_6 <- file_6[, !colnames(file_6) %in% col_drop]
gc()
```

```{r}
file_7 <- read_parquet("/Users/ishaan/Downloads/part-7.parquet")
file_7 <- file_7[, !colnames(file_7) %in% col_drop]
gc()
```

```{r}
file_8 <- read_parquet("/Users/ishaan/Downloads/part-8.parquet")
file_8 <- file_8[, !colnames(file_8) %in% col_drop]
gc()
```

```{r}
file_9 <- read_parquet("/Users/ishaan/Downloads/part-9.parquet")
file_9 <- file_9[, !colnames(file_9) %in% col_drop]
gc()
```

```{r}
file_10 <- read_parquet("/Users/ishaan/Downloads/part-10.parquet")
file_10 <- file_10[, !colnames(file_10) %in% col_drop]
head(file_10)
gc()
```

```{r}
colnames(file_10)[colSums(is.na(file_10)) > 0]
```

```{r}
colnames(file_10)[sapply(file_10, function(col) {
  is.character(col) && any(col == "None", na.rm = TRUE)
})]
```

**Replacing "None" by 0 to ease conversion to integer.**
```{r}
file_1$in.pv_system_size[file_1$in.pv_system_size == "None"] <- 0
file_2$in.pv_system_size[file_2$in.pv_system_size == "None"] <- 0
file_3$in.pv_system_size[file_3$in.pv_system_size == "None"] <- 0
file_4$in.pv_system_size[file_4$in.pv_system_size == "None"] <- 0
file_5$in.pv_system_size[file_5$in.pv_system_size == "None"] <- 0
file_6$in.pv_system_size[file_6$in.pv_system_size == "None"] <- 0
file_7$in.pv_system_size[file_7$in.pv_system_size == "None"] <- 0
file_8$in.pv_system_size[file_8$in.pv_system_size == "None"] <- 0
file_9$in.pv_system_size[file_9$in.pv_system_size == "None"] <- 0
file_10$in.pv_system_size[file_10$in.pv_system_size == "None"] <- 0
```

**Function to remove character values from a column and cast it to int**

```{r}
clean_and_convert_to_integer <- function(data, columns_to_clean) {
  # `columns_to_clean` should be a named list where the name is the column name
  # and the value is a vector of characters to remove for that column
  
  for (column_name in names(columns_to_clean)) {
    # Check if the column exists in the data frame
    if (!column_name %in% colnames(data)) {
      warning(paste("Column", column_name, "not found in the data frame. Skipping."))
      next
    }
    
    # Ensure the column is character or numeric
    if (!is.character(data[[column_name]]) && !is.numeric(data[[column_name]])) {
      warning(paste("Column", column_name, "is not character or numeric. Skipping."))
      next
    }
    
    # Get the list of characters to remove for the column
    characters_to_remove <- columns_to_clean[[column_name]]
    
    # Replace each specified character in the column
    for (char in characters_to_remove) {
      data[[column_name]] <- gsub(char, "", data[[column_name]])
    }
   
    # Convert the column to integer
    data[[column_name]] <- as.integer(data[[column_name]])
  }
  
  # Return the modified data frame
  return(data)
}

```


**Defining the columns to clean**
```{r}
# Define the columns and their respective characters to remove
int_columns_to_clean <- list(
  "in.bathroom_spot_vent_hour" = c("Hour"),
  "in.cooling_setpoint" = c("F"), 
  "in.cooling_setpoint_offset_magnitude" = c("F"),
  "in.heating_setpoint" = c("F"), 
  "in.heating_setpoint_offset_magnitude" = c("F"), 
  "in.plug_load_diversity" = c("%"), 
  "in.range_spot_vent_hour" = c("Hour"), 
  "in.infiltration" = c(" ACH50"), 
  "in.occupants" = c("+"),
  "in.pv_system_size" = c(".0 kWDC")
)
```

**Function to convert character variables with 2 values to a factor**

```{r}
convert_character_columns_to_factors <- function(data, columns_to_convert, custom_levels = NULL) {
  # `columns_to_convert` should be a vector of column names to convert
  # `custom_levels` is a named list specifying custom levels for certain columns
  
  for (column_name in columns_to_convert) {
    # Check if the column exists in the data frame
    if (!column_name %in% colnames(data)) {
      warning(paste("Column", column_name, "not found in the data frame. Skipping."))
      next
    }
    
    # Ensure the column is character
    if (!is.character(data[[column_name]])) {
      warning(paste("Column", column_name, "is not a character column. Skipping."))
      next
    }
    
    # Get the unique values in the column
    unique_values <- unique(data[[column_name]])
    
    # Check if the column has exactly two unique values
    if (length(unique_values) != 2) {
      warning(paste("Column", column_name, "does not have exactly two unique values. Skipping."))
      next
    }
    
    # Determine the levels for the factor
    if (!is.null(custom_levels) && column_name %in% names(custom_levels)) {
      levels <- custom_levels[[column_name]]
    } else {
      levels <- sort(unique_values)  # Default to alphabetical order if no custom levels provided
    }
    
    # Convert the column to a factor with specified levels
    data[[column_name]] <- factor(data[[column_name]], levels = levels, labels = c(0, 1), ordered = TRUE)
  }
  
  # Return the modified data frame
  return(data)
}


```

**Defining the columns to clean**
```{r}
# Define the columns to convert
binary_columns_to_convert <- c( "in.cooling_setpoint_has_offset", "in.has_pv", "in.heating_setpoint_has_offset", 
                         "in.hvac_has_ducts", "in.hvac_has_zonal_electric_heating", "in.misc_gas_fireplace",
                         "in.misc_gas_grill", "in.misc_gas_lighting", "in.misc_pool", "in.misc_pool_pump", 
                         "in.misc_well_pump", "in.tenure", "in.vacancy_status")

# Define custom levels
binary_custom_levels <- list(
  in.cooling_setpoint_has_offset = c("No", "Yes"),
  in.has_pv = c("No", "Yes"), 
  in.heating_setpoint_has_offset = c("No", "Yes"),
  in.hvac_has_ducts = c("No", "Yes"),
  in.hvac_has_zonal_electric_heating = c("No", "Yes"),
  in.misc_gas_fireplace = c("None", "Gas Fireplace"), 
  in.misc_gas_grill = c("None", "Gas Grill"), 
  in.misc_gas_lighting = c("None", "Gas Lighting"), 
  in.misc_pool = c("None", "1.0 HP Pump"),
  in.misc_well_pump = c("None", "Typical Efficiency"), 
  in.tenure = c("Renter", "Owner"), 
  in.vacancy_status = c("Vacant", "Occupied")
)
```

**Using gsub() to convert all the fuel columns to Electricity and "Other Fuel"**

```{r}
#File 1
file_1$in.clothes_dryer <- gsub("Gas", "Propane", file_1$in.clothes_dryer)
file_1$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_1$in.clothes_dryer)
file_1$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_1$in.clothes_dryer)

file_1$in.cooking_range <- gsub("Gas", "Propane", file_1$in.cooking_range)
file_1$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_1$in.cooking_range)
file_1$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_1$in.cooking_range)

file_1$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_1$in.heating_fuel)
file_1$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_1$in.heating_fuel)
file_1$in.heating_fuel <- gsub("Propane", "Other Fuel", file_1$in.heating_fuel)

#File 2
file_2$in.clothes_dryer <- gsub("Gas", "Propane", file_2$in.clothes_dryer)
file_2$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_2$in.clothes_dryer)
file_2$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_2$in.clothes_dryer)

file_2$in.cooking_range <- gsub("Gas", "Propane", file_2$in.cooking_range)
file_2$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_2$in.cooking_range)
file_2$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_2$in.cooking_range)

file_2$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_2$in.heating_fuel)
file_2$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_2$in.heating_fuel)
file_2$in.heating_fuel <- gsub("Propane", "Other Fuel", file_2$in.heating_fuel)

#File 3
file_3$in.clothes_dryer <- gsub("Gas", "Propane", file_3$in.clothes_dryer)
file_3$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_3$in.clothes_dryer)
file_3$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_3$in.clothes_dryer)

file_3$in.cooking_range <- gsub("Gas", "Propane", file_3$in.cooking_range)
file_3$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_3$in.cooking_range)
file_3$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_3$in.cooking_range)

file_3$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_3$in.heating_fuel)
file_3$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_3$in.heating_fuel)
file_3$in.heating_fuel <- gsub("Propane", "Other Fuel", file_3$in.heating_fuel)

#File 4
file_4$in.clothes_dryer <- gsub("Gas", "Propane", file_4$in.clothes_dryer)
file_4$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_4$in.clothes_dryer)
file_4$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_4$in.clothes_dryer)

file_4$in.cooking_range <- gsub("Gas", "Propane", file_4$in.cooking_range)
file_4$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_4$in.cooking_range)
file_4$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_4$in.cooking_range)

file_4$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_4$in.heating_fuel)
file_4$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_4$in.heating_fuel)
file_4$in.heating_fuel <- gsub("Propane", "Other Fuel", file_4$in.heating_fuel)

#File 5
file_5$in.clothes_dryer <- gsub("Gas", "Propane", file_5$in.clothes_dryer)
file_5$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_5$in.clothes_dryer)
file_5$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_5$in.clothes_dryer)

file_5$in.cooking_range <- gsub("Gas", "Propane", file_5$in.cooking_range)
file_5$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_5$in.cooking_range)
file_5$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_5$in.cooking_range)

file_5$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_5$in.heating_fuel)
file_5$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_5$in.heating_fuel)
file_5$in.heating_fuel <- gsub("Propane", "Other Fuel", file_5$in.heating_fuel)

#File 6
file_6$in.clothes_dryer <- gsub("Gas", "Propane", file_6$in.clothes_dryer)
file_6$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_6$in.clothes_dryer)
file_6$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_6$in.clothes_dryer)

file_6$in.cooking_range <- gsub("Gas", "Propane", file_6$in.cooking_range)
file_6$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_6$in.cooking_range)
file_6$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_6$in.cooking_range)

file_6$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_6$in.heating_fuel)
file_6$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_6$in.heating_fuel)
file_6$in.heating_fuel <- gsub("Propane", "Other Fuel", file_6$in.heating_fuel)

#File 7
file_7$in.clothes_dryer <- gsub("Gas", "Propane", file_7$in.clothes_dryer)
file_7$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_7$in.clothes_dryer)
file_7$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_7$in.clothes_dryer)

file_7$in.cooking_range <- gsub("Gas", "Propane", file_7$in.cooking_range)
file_7$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_7$in.cooking_range)
file_7$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_7$in.cooking_range)

file_7$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_7$in.heating_fuel)
file_7$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_7$in.heating_fuel)
file_7$in.heating_fuel <- gsub("Propane", "Other Fuel", file_7$in.heating_fuel)

#File 8
file_8$in.clothes_dryer <- gsub("Gas", "Propane", file_8$in.clothes_dryer)
file_8$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_8$in.clothes_dryer)
file_8$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_8$in.clothes_dryer)

file_8$in.cooking_range <- gsub("Gas", "Propane", file_8$in.cooking_range)
file_8$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_8$in.cooking_range)
file_8$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_8$in.cooking_range)

file_8$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_8$in.heating_fuel)
file_8$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_8$in.heating_fuel)
file_8$in.heating_fuel <- gsub("Propane", "Other Fuel", file_8$in.heating_fuel)

#File 9
file_9$in.clothes_dryer <- gsub("Gas", "Propane", file_9$in.clothes_dryer)
file_9$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_9$in.clothes_dryer)
file_9$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_9$in.clothes_dryer)

file_9$in.cooking_range <- gsub("Gas", "Propane", file_9$in.cooking_range)
file_9$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_9$in.cooking_range)
file_9$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_9$in.cooking_range)

file_9$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_9$in.heating_fuel)
file_9$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_9$in.heating_fuel)
file_9$in.heating_fuel <- gsub("Propane", "Other Fuel", file_9$in.heating_fuel)

#File 10
file_10$in.clothes_dryer <- gsub("Gas", "Propane", file_10$in.clothes_dryer)
file_10$in.clothes_dryer <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_10$in.clothes_dryer)
file_10$in.clothes_dryer <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_10$in.clothes_dryer)

file_10$in.cooking_range <- gsub("Gas", "Propane", file_10$in.cooking_range)
file_10$in.cooking_range <- gsub("Propane, 80% Usage", "Propane, 120% Usage", file_10$in.cooking_range)
file_10$in.cooking_range <- gsub("Propane, 100% Usage", "Propane, 120% Usage", file_10$in.cooking_range)

file_10$in.heating_fuel <- gsub("Natural Gas", "Other Fuel", file_10$in.heating_fuel)
file_10$in.heating_fuel <- gsub("Fuel Oil", "Other Fuel", file_10$in.heating_fuel)
file_10$in.heating_fuel <- gsub("Propane", "Other Fuel", file_10$in.heating_fuel)


```

**Income column has a discrepancy. One value within it says <10000 but all values are above 10000 anyways. Should it be <100000 or  >10000 instead?**

**Converting "None" to NA in these columns as it is not an allowable value**

```{r}
#These columns having "None" makes no sense, thus they're being converted to NA before the column is cast as a factor

file_1$in.ducts[file_1$in.ducts == "None"] <- NA
file_1$in.geometry_wall_exterior_finish[file_1$in.geometry_wall_exterior_finish == "None"] <- NA
file_1$in.heating_fuel[file_1$in.heating_fuel == "None"] <- NA
file_1$in.hvac_heating_efficiency[file_1$in.hvac_heating_efficiency == "Other"] <- NA
file_1$in.insulation_ceiling[file_1$in.insulation_ceiling == "None"] <- NA
file_1$in.insulation_floor[file_1$in.insulation_floor == "None"] <- NA
file_1$in.insulation_foundation_wall[file_1$in.insulation_foundation_wall == "None"] <- NA
file_1$in.insulation_rim_joist[file_1$in.insulation_rim_joist == "None"] <- NA
file_1$in.insulation_slab[file_1$in.insulation_slab == "NA"] <- NA
file_1$in.water_heater_efficiency[file_1$in.water_heater_efficiency == "Other Fuel"] <- NA
file_1$in.water_heater_fuel[file_1$in.water_heater_fuel == "Other Fuel"] <- NA

```

**Function to convert all columns with a valid "None" entry to factor with levels starting from 0**

```{r}
convert_with_none <- function(data, columns_to_convert, custom_levels) {
  # `columns_to_convert` is a vector of column names
  # `custom_levels` is a named list where each column has its sorted levels
  
  for (column_name in columns_to_convert) {
    # Check if the column exists in the data frame
    if (!column_name %in% colnames(data)) {
      warning(paste("Column", column_name, "not found in the data frame. Skipping."))
      next
    }
    
    # Ensure the column is character
    if (!is.character(data[[column_name]])) {
      warning(paste("Column", column_name, "is not a character column. Skipping."))
      next
    }
    
    # Get the custom levels for the column
    if (!column_name %in% names(custom_levels)) {
      warning(paste("No custom levels provided for column", column_name, "Skipping."))
      next
    }
    
    levels <- custom_levels[[column_name]]
    
    # Convert to factor with levels starting from 0
    data[[column_name]] <- factor(data[[column_name]], levels = levels, labels = seq_along(levels) - 1)
  }
  return(data)
}

```

**All columns with valid NA values and their respective customized levels**

```{r}
valid_none_columns_to_convert <- c("in.ceiling_fan", "in.clothes_washer", "in.dishwasher", "in.geometry_attic_type", "in.geometry_garage",
                                   "in.hvac_cooling_efficiency","in.hvac_cooling_partial_space_conditioning","in.hvac_cooling_type",
                                   "in.hvac_heating_efficiency","in.hvac_heating_type","in.hvac_heating_type_and_fuel",
                                   "in.misc_extra_refrigerator","in.misc_freezer", "in.misc_hot_tub_spa", "in.misc_pool_heater",
                                   "in.refrigerator")
  
valid_none_custom_levels <- list(
  in.ceiling_fan = c("None", "Standard Efficiency, No Usage", "Standard Efficiency"),
  in.clothes_washer = c("None", "EnergyStar, 80% Usage", "EnergyStar, 100% Usage", "EnergyStar, 120% Usage",  "Standard, 80%
                        Usage", "Standard, 100% Usage", "Standard, 120% Usage"), 
  in.dishwasher = c("None", "290 Rated kWh, 80% Usage", "318 Rated kWh, 80% Usage", "290 Rated kWh, 100% Usage", "318 Rated kWh, 100% Usage",
                    "290 Rated kWh, 120% Usage", "318 Rated kWh, 120% Usage"),
  in.geometry_attic_type = c("None", "Finished Attic or Cathedral Ceilings", "Vented Attic"),
  in.geometry_garage = c("None", "1 Car", "2 Car", "3 Car"), 
  in.hvac_cooling_efficiency = c("None", "AC, SEER 8", "AC, SEER 10", "AC, SEER 13", "AC, SEER 15", "Room AC, EER 8.5", 
                                 "Room AC, EER 9.8", "Room AC, EER 10.7", "Room AC, EER 12.0", "Heat Pump", "Shared Cooling"),
  in.hvac_cooling_partial_space_conditioning = c("None", "<10% Conditioned", "20% Conditioned", "40% Conditioned", "60%
                                                 Conditioned", "80% Conditioned", "100% Conditioned"),
  in.hvac_cooling_type = c("None", "Room AC", "Heating Pump", "Cantral AC"),
  in.hvac_heating_efficiency = c("None", "Shared Heating", "Fuel Furnace, 92.5% AFUE", "Fuel Wall/Floor Furnace, 92.5% AFUE", "Fuel
                                 Furnace, 80%AFUE","Fuel Wall/Floor Furnace, 80% AFUE", "Fuel Boiler, 90% AFUE", "Fuel Furnace, 76%
                                 AFUE","Fuel Boiler, 80% AFUE", "Fuel Wall/Floor Furnace, 68% AFUE", "ASHP, SEER 15, 8.5
                                 HSPF","ASHP, SEER 13, 7.7 HSPF", "ASHP, SEER 10, 6.2 HSPF", "Electric Furnace, 100% AFUE",
                                 "Electric Boiler, 100% AFUE", "Electric Baseboard, 100% Efficiency", "Electric Wall Furnace, 100%
                                 AFUE"),
  in.hvac_heating_type = c("None", "Ducted Heat Pump", "Ducted Heating", "Non-Ducted Heating"), 
  in.hvac_heating_type_and_fuel = c("None", "Fuel Oil Shared Heating", "Natural Gas Shared Heating", "Propane Shared Heating",
                                    "Fuel Oil Fuel Wall/Floor Furnace", "Natural Gas Fuel Wall/Floor Furnace", "Propane Fuel
                                    Wall/Floor Furnace", "Fuel Oil Fuel Furnace", "Natural Gas Fuel Furnace", "Propane Fuel
                                    Furnace", "Fuel Oil Fuel Boiler", "Natural Gas Fuel Boiler", "Propane Fuel Boiler",
                                    "Electricity Shared Heating", "Electricity Electric Wall Furnace", "Electricity Electric
                                    Furnace", "Electricity Electric Boiler", "Electricity Baseboard", "Electricity ASHP"),
  in.misc_extra_refrigerator = c("None", "EF 19.9", "EF 17.6", "EF 15.9", "EF 10.5", "EF 10.2", "EF 6.7"),
  in.misc_freezer = c("None", "National Average", "EF 12"),
  in.misc_hot_tub_spa = c("None", "Gas", "Electricity"),
  in.misc_pool_heater = c("None", "Solar", "Gas", "Electric"),
  in.refrigerator = c("None", "EF 19.9, 100% Usage", "EF 17.6, 100% Usage", "EF 15.9, 100% Usage", "EF 10.5, 100% Usage", "EF 10.2, 100%
                      Usage", "EF 6.7, 100% Usage"))
```

**Function to convert all columns with invalid "None" entries to factor with levels starting from 1**

```{r}
convert_without_none <- function(data, columns_to_convert, custom_levels) {
  # `columns_to_convert` is a vector of column names
  # `custom_levels` is a named list where each column has its sorted levels
  
  for (column_name in columns_to_convert) {
    # Check if the column exists in the data frame
    if (!column_name %in% colnames(data)) {
      warning(paste("Column", column_name, "not found in the data frame. Skipping."))
      next
    }
    
    # Ensure the column is character
    if (!is.character(data[[column_name]])) {
      warning(paste("Column", column_name, "is not a character column. Skipping."))
      next
    }
    
    # Get the custom levels for the column
    if (!column_name %in% names(custom_levels)) {
      warning(paste("No custom levels provided for column", column_name, "Skipping."))
      next
    }
    
    levels <- custom_levels[[column_name]]
    
    # Convert to factor with levels starting from 1
    data[[column_name]] <- factor(data[[column_name]], levels = levels, labels = seq_along(levels))
  }
  return(data)
}

```

**All columns with invalid NA values and their respective customized levels**

```{r}
invalid_none_columns_to_convert <- c("in.building_america_climate_zone", "in.clothes_dryer",  "in.cooking_range", "in.ducts",
                                     "in.geometry_floor_area", "in.geometry_foundation_type", "in.geometry_wall_exterior_finish",
                                     "in.geometry_wall_type", "in.heating_fuel", "in.hot_water_fixtures", "in.insulation_ceiling",
                                     "in.insulation_floor","in.insulation_foundation_wall", "in.insulation_rim_joist",
                                     "in.insulation_roof","in.insulation_slab", "in.insulation_wall", "in.lighting",
                                     "in.roof_material","in.usage_level", "in.water_heater_efficiency", "in.water_heater_fuel",
                                     "in.window_areas", "in.windows")



invalid_none_custom_levels <- list(
  in.building_america_climate_zone = c("Marine", "Mixed-Humid", "Mixed-Dry", "Hot-Humid", "Hot-Dry", "Cold", "Very Cold"),
  in.clothes_dryer = c("Propane, 120% Usage", "Electric, 80% Usage", "Electric, 100% Usage", "Electric, 120% Usage"),
  in.cooking_range = c("Propane, 120% Usage", "Electric, 80% Usage", "Electric, 100% Usage", "Electric, 120% Usage"),
  in.ducts = c("0% Leakage, Uninsulated", "10% Leakage, R-8", "10% Leakage, R-6,", "10% Leakage, R-4", "20% Leakage, R-8", "30%
               Leakage, R-8", "20% Leakage, R-6", "20% Leakage, R-4", "30% Leakage, R-6", "10% Leakage, Uninsulated", "20% Leakage,
               Uninsulated", "30% Leakage, R-4", "30% Leakage, Uninsulated"),
  in.geometry_floor_area = c("0-499", "500-749", "750-999", "1000-1499", "1500-1999", "2000-2499", "2500-2999", "3000-3999",
                             "4000+"),
  in.geometry_foundation_type = c("Ambient", "Slab", "Unvented Crawlspace", "Unheated Basement", "Vented Crawlspace", "Heated Basement"),
  in.geometry_wall_exterior_finish = c("Aluminum, Light", "Vinyl, Light", "Stucco, Light", "Brick, Light", "Fiber-Cement, Light",
                                       "Stucco, Medium/Dark", "Brick, Medium/Dark", "Wood, Medium/Dark", "Shingle, Composition,
                                       Medium", "Shingle, Asbestos, Medium"),
  in.geometry_wall_type = c("Concrete", "Steel Frame", "Brick", "Wood Frame"),
  in.heating_fuel = c("Other Fuel", "Electricity"),
  in.hot_water_fixtures = c("50% Usage", "100% Usage", "200% Usage"),
  in.insulation_ceiling = c("R-49", "R-38", "R-30", "R-19", "R-13", "R-7", "Uninsulated"),
  in.insulation_floor = c("Ceiling R-30", "Ceiling R-19", "Ceiling R-13", "Uninsulated"),
  in.insulation_foundation_wall = c("Wall R-15, Exterior", "Wall R-10, Exterior", "Wall R-5, Exterior", "Uninsulated"),
  in.insulation_rim_joist = c("Uninsulated", "R-5, Exterior", "R-10, Exterior", "R-15, Exterior"),
  in.insulation_roof = c("Finished, R-49", "Finished, R-38", "Finished, R-30", "Finished, R-19", "Finished, R-13", "Finished, R-7",
                         "Finished, Uninsulated", "Unfinished, Uninsulated"),
  in.insulation_slab = c("2ft R10 Perimeter, Vertical", "2ft R10 Under, Horizontal", "2ft R5 Perimeter, Vertical", "2ft R5 Under,
                         Horizontal", "Uninsulated"),
  in.insulation_wall = c("Brick, 12-in, 3-wythe, R-19", "CMU, 6-in Hollow, R-19", "Wood Stud, R-19", "Brick, 12-in, 3-wythe, R-15", "CMU, 6-in
                         Hollow, R-15", "Wood Stud, R-15", "Brick, 12-in, 3-wythe, R-11", "CMU, 6-in Hollow, R-11", "Wood Stud, R-11", "Brick,
                         12-in, 3-wythe, R-7", "CMU, 6-in Hollow, R-7", "Wood Stud, R-7", "Brick, 12-in, 3-wythe, Uninsulated", "CMU, 6-in
                         Hollow, Uninsulated", "Wood Stud, Uninsulated"),
  in.lighting = c("100% LED", "100% CFL", "100% Incandescent"),
  in.roof_material = c("Tile, Concrete", "Tile, Clay or Ceramic", "Slate", "Wood Shingles", "Composition Shingles", "Asphalt Shingles,
                       Medium", "Metal, Dark"),
  in.usage_level = c("Low", "Medium", "High"), 
  in.water_heater_efficiency = c("Natural Gas Tankless", "Propane Tankless", "FIXME Fuel Oil Indirect", "Fuel Oil Indirect", "Natural Gas
                                 Premium", "Propane Premium", "Fuel Oil Premium", "Natural Gas Standard", "Propane Standard", "Fuel Oil
                                 Standard", "Electric Tankless", "Electric Premium", "Electric Standard", "Electric Heat Pump, 80 gal"),
  in.water_heater_fuel = c("Natural Gas", "Fuel Oil", "Propane", "Electricity"),
  in.window_areas = c("F6 B6 L6 R6", "F9 B9 L9 R9", "F12 B12 L12 R12", "F15 B15 L15 R15", "F18 B18 L18 R18", "F30 B30 L30 R30"),
  in.windows = c("Triple, Low-E, Non-metal, Air, L-Gain", "Double, Low-E, Non-metal, Air, M-Gain", "Double, Clear, Non-metal, Air, Exterior
                 Clear Storm", "Double, Clear, Non-metal, Air", "Double, Clear, Metal, Air, Exterior Clear Storm", "Double, Clear, Metal,
                 Air", "Single, Clear, Non-metal", "Single, Clear, Non-metal, Exterior Clear Storm", "Single, Clear, Metal", "Single, Clear,
                 Metal, Exterior Clear Storm"))
```


**Finally, convert the setpoint columns from string to int by doing the appropriate conversion**

```{r}
convert_cooling_setpoint_offset_period <- function(data, column_name) {
  # Define the mapping of characters to integers
  mapping <- c(
    "Day Setup" = 8,
    "Day Setup +1h" = 9,
    "Day Setup +2h" = 10,
    "Day Setup +3h" = 11,
    "Day Setup +4h" = 12,
    "Day Setup +5h" = 13,
    "Day Setup -1h" = 7,
    "Day Setup -2h" = 6,
    "Day Setup -3h" = 5,
    "Day Setup -4h" = 4,
    "Day Setup -5h" = 3,
    "Day Setup and Night Setback" = 17,
    "Day Setup and Night Setback +1h" = 18,
    "Day Setup and Night Setback +2h" = 19,
    "Day Setup and Night Setback +3h" = 20,
    "Day Setup and Night Setback +4h" = 21,
    "Day Setup and Night Setback +5h" = 22,
    "Day Setup and Night Setback -1h" = 16,
    "Day Setup and Night Setback -2h" = 15,
    "Day Setup and Night Setback -3h" = 14,
    "Day Setup and Night Setback -4h" = 13,
    "Day Setup and Night Setback -5h" = 12,
    "Day and Night Setup" = 17,
    "Day and Night Setup +1h" = 18,
    "Day and Night Setup +2h" = 19,
    "Day and Night Setup +3h" = 20,
    "Day and Night Setup +4h" = 21,
    "Day and Night Setup +5h" = 22,
    "Day and Night Setup -1h" = 16,
    "Day and Night Setup -2h" = 15,
    "Day and Night Setup -3h" = 14,
    "Day and Night Setup -4h" = 13,
    "Day and Night Setup -5h" = 12,
    "Night Setback" = 9,
    "Night Setback +1h" = 10,
    "Night Setback +2h" = 11,
    "Night Setback +3h" = 12,
    "Night Setback +4h" = 13,
    "Night Setback +5h" = 14,
    "Night Setback -1h" = 8,
    "Night Setback -2h" = 7,
    "Night Setback -3h" = 6,
    "Night Setback -4h" = 5,
    "Night Setback -5h" = 4,
    "Night Setup" = 9,
    "Night Setup +1h" = 10,
    "Night Setup +2h" = 11,
    "Night Setup +3h" = 12,
    "Night Setup +4h" = 13,
    "Night Setup +5h" = 14,
    "Night Setup -1h" = 8,
    "Night Setup -2h" = 7,
    "Night Setup -3h" = 6,
    "Night Setup -4h" = 5,
    "Night Setup -5h" = 4,
    "None" = 0
  )
  
  # Check if the column exists in the data frame
  if (!column_name %in% colnames(data)) {
    stop(paste("Column", column_name, "not found in the data frame."))
  }
  
  # Map the column values to integers using the mapping
  data[[column_name]] <- sapply(data[[column_name]], function(value) {
    if (value %in% names(mapping)) {
      return(mapping[value])
    } else {
      warning(paste("Unrecognized value:", value))
      return(NA)  # Return NA for unrecognized values
    }
  })
  
  # Ensure the column is numeric
  data[[column_name]] <- as.numeric(data[[column_name]])
  
  return(data)
}
```

```{r}
convert_heating_setpoint_offset_period <- function(data, column_name) {
  # Define the mapping of characters to integers
  mapping <- c(
    "Day" = 8,
    "Day +1h" = 9,
    "Day +2h" = 10,
    "Day +3h" = 11,
    "Day +4h" = 12,
    "Day +5h" = 13,
    "Day -1h" = 7,
    "Day -2h" = 6,
    "Day -3h" = 5,
    "Day -4h" = 4,
    "Day -5h" = 3,
    "Day and Night" = 17,
    "Day and Night +1h" = 18,
    "Day and Night +2h" = 19,
    "Day and Night +3h" = 20,
    "Day and Night +4h" = 21,
    "Day and Night +5h" = 22,
    "Day and Night -1h" = 16,
    "Day and Night -2h" = 15,
    "Day and Night -3h" = 14,
    "Day and Night -4h" = 13,
    "Day and Night -5h" = 12,
    "Night" = 9,
    "Night +1h" = 10,
    "Night +2h" = 11,
    "Night +3h" = 12,
    "Night +4h" = 13,
    "Night +5h" = 14,
    "Night -1h" = 8,
    "Night -2h" = 7,
    "Night -3h" = 6,
    "Night -4h" = 5,
    "Night -5h" = 4,
    "None" = 0
  )

  
  # Check if the column exists in the data frame
  if (!column_name %in% colnames(data)) {
    stop(paste("Column", column_name, "not found in the data frame."))
  }
  
  # Map the column values to integers using the mapping
  data[[column_name]] <- sapply(data[[column_name]], function(value) {
    if (value %in% names(mapping)) {
      return(mapping[value])
    } else {
      warning(paste("Unrecognized value:", value))
      return(NA)  # Return NA for unrecognized values
    }
  })
  
  # Ensure the column is numeric
  data[[column_name]] <- as.numeric(data[[column_name]])
  
  return(data)
}
```



**Applying all the functions to each file in order**

```{r}
#File 1
file_1 <- clean_and_convert_to_integer(file_1, int_columns_to_clean)
file_1 <- convert_character_columns_to_factors(file_1, binary_columns_to_convert, binary_custom_levels)
file_1 <- convert_with_none(file_1, valid_none_columns_to_convert, valid_none_custom_levels)
file_1 <- convert_without_none(file_1, invalid_none_columns_to_convert, invalid_none_custom_levels)
file_1 <- convert_cooling_setpoint_offset_period(file_1, "in.cooling_setpoint_offset_period") #ONLY RUN ONCE
file_1 <- convert_heating_setpoint_offset_period(file_1, "in.heating_setpoint_offset_period") #ONLY RUN ONCE
```

```{r}
#File 2
file_2 <- clean_and_convert_to_integer(file_2, int_columns_to_clean)
file_2 <- convert_character_columns_to_factors(file_2, binary_columns_to_convert, binary_custom_levels)
file_2 <- convert_with_none(file_2, valid_none_columns_to_convert, valid_none_custom_levels)
file_2 <- convert_without_none(file_2, invalid_none_columns_to_convert, invalid_none_custom_levels)
```

```{r}
#File 3
file_3 <- clean_and_convert_to_integer(file_3, int_columns_to_clean)
file_3 <- convert_character_columns_to_factors(file_2, binary_columns_to_convert, binary_custom_levels)
file_3<- convert_with_none(file_3, valid_none_columns_to_convert, valid_none_custom_levels)
file_3 <- convert_without_none(file_3, invalid_none_columns_to_convert, invalid_none_custom_levels)
```


```{r}
#File 4
file_4 <- clean_and_convert_to_integer(file_4, int_columns_to_clean)
file_4 <- convert_character_columns_to_factors(file_4, binary_columns_to_convert, binary_custom_levels)
file_4 <- convert_with_none(file_4, valid_none_columns_to_convert, valid_none_custom_levels)
file_4 <- convert_without_none(file_4, invalid_none_columns_to_convert, invalid_none_custom_levels)
```


```{r}
#File 5
file_5 <- clean_and_convert_to_integer(file_5, int_columns_to_clean)
file_5 <- convert_character_columns_to_factors(file_5, binary_columns_to_convert, binary_custom_levels)
file_5 <- convert_with_none(file_5, valid_none_columns_to_convert, valid_none_custom_levels)
file_5 <- convert_without_none(file_5, invalid_none_columns_to_convert, invalid_none_custom_levels)
```

```{r}
#File 6
file_6 <- clean_and_convert_to_integer(file_6, int_columns_to_clean)
file_6 <- convert_character_columns_to_factors(file_6, binary_columns_to_convert, binary_custom_levels)
file_6 <- convert_with_none(file_6, valid_none_columns_to_convert, valid_none_custom_levels)
file_6 <- convert_without_none(file_6, invalid_none_columns_to_convert, invalid_none_custom_levels)
```

```{r}
#File 7
file_7 <- clean_and_convert_to_integer(file_7, int_columns_to_clean)
file_7 <- convert_character_columns_to_factors(file_7, binary_columns_to_convert, binary_custom_levels)
file_7 <- convert_with_none(file_7, valid_none_columns_to_convert, valid_none_custom_levels)
file_7 <- convert_without_none(file_7, invalid_none_columns_to_convert, invalid_none_custom_levels)
```

```{r}
#File 8
file_9 <- clean_and_convert_to_integer(file_9, int_columns_to_clean)
file_9 <- convert_character_columns_to_factors(file_9, binary_columns_to_convert, binary_custom_levels)
file_9 <- convert_with_none(file_9, valid_none_columns_to_convert, valid_none_custom_levels)
file_9 <- convert_without_none(file_9, invalid_none_columns_to_convert, invalid_none_custom_levels)
```

```{r}
#File 10
file_10 <- clean_and_convert_to_integer(file_10, int_columns_to_clean)
file_10 <- convert_character_columns_to_factors(file_10, binary_columns_to_convert, binary_custom_levels)
file_10 <- convert_with_none(file_10, valid_none_columns_to_convert, valid_none_custom_levels)
file_10 <- convert_without_none(file_10, invalid_none_columns_to_convert, invalid_none_custom_levels)
```

```{r}
which(colSums(is.na(file_1)) > 0)
```
```{r}
weather_columns <- c("Dry Bulb Temperature [°C]", "Relative Humidity [%]", "Wind Speed [m/s]", "Wind Direction [Deg]", "Global Horizontal Radiation [W/m2]", "Direct Normal Radiation [W/m2]", "Diffuse Horizontal Radiation [W/m2]")

file_1[apply(file_1[weather_columns], 1, function(row) all(is.na(row))), ]

```

